---
title: "Polygon Analysis"
author: "Jackson Krebsbach"
date: "9/28/2021"
output:
html_document: default
pdf_document: default
---

## Set up

```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=normalizePath("..")) #set working directory to the project directory
```

```{r}
library(tidyverse)
library(raster)
library(ggplot2)
require(gridExtra)
library(cowplot)
library(sf)
library(dplyr)
library(RStoolbox)
library(jsonlite)
source("code/helpers/veggie_functions.R")
```

## Read in the data
```{r include = FALSE}
poly_pixels <- readRDS('clean_data/rdata/pixel_values/features/poly_pixels.rds')

polys_sf <- readRDS('./clean_data/rdata/pixel_values/features/polys_sf.rds') %>%
  st_make_valid() %>%
  as_tibble()

val_pred <- readRDS("./clean_data/rdata/pixel_values/predictions/all_pixels_val_pixels.rds")

val_pred <-  val_pred %>%   
    left_join(polys_sf) %>%
    dplyr::select(c(label, .pred_class, poly_num, quadrat, area))
```


## Reading in data
<!-- ## Filter Live Veg -->
```{r, include = FALSE}
subset <- val_pred 
subset %>% filter(label == 'live vegetation') -> live_vegetation_data
subset %>% filter(label == 'dead vegetation') -> dead_vegetation_data
subset %>% filter(label == 'sand') -> sand_data
```


## Plotting Bar Charts for live vetation, dead vegetation, and sand
```{r fig.height = 8, fig.width = 9, fig.align = "center", dpi=350, echo=FALSE}
dead_vegetation_data %>%
  mutate(area_class = cut(area, seq(0, max(area) + 50, 50 ), include.lowest = TRUE, right = TRUE ))  -> dead_vegetation

dead_vegetation %>%
  mutate(poly_id = paste(poly_num, quadrat, sep = "_")) %>%
  group_by(area_class, poly_id) %>%
  summarise(area = first(area), area_class = first(area_class)) %>%
  summarize(area = sum(area), poly_count = n() ) -> dead_vegetation_summary 

sum_plt <-  ggplot(dead_vegetation_summary, aes(x = area_class, y = area, label = poly_count)) +
  geom_col()  +
  geom_text(nudge_y = 250, size = 3.5) +
  theme(axis.text.x = element_text(angle = 75, hjust=1,size = 9))
  
class_plt <- ggplot(dead_vegetation, aes(fill=.pred_class, y = 1, x=area_class)) + 
   geom_bar(position="fill", stat="identity") +  
   theme(axis.text.x = element_text(angle = 75, hjust=1,size = 10)) +
  theme(legend.position = "bottom")

grid.arrange(sum_plt, class_plt, ncol=1,top = "Dead Vegetation") 
```


### Live vegetation
```{r fig.height = 8, fig.width = 9, fig.align = "center", dpi=350, echo=FALSE}
live_vegetation_data %>%
  drop_na() %>%
  mutate(area_class = cut(area, seq(0, max(area) + 50, 50 ), include.lowest = TRUE, right = TRUE ))  -> live_vegetation

live_vegetation %>%
  mutate(poly_id = paste(poly_num, quadrat, sep = "_")) %>%
  group_by(area_class, poly_id) %>%
  summarise(area = first(area), area_class = first(area_class)) %>%
  summarize(area = sum(area), poly_count = n() ) -> live_vegetation_summary 

sum_plt <-  ggplot(live_vegetation_summary, aes(x = area_class, y = area, label = poly_count)) +
  geom_col()  +
  geom_text(nudge_y = 250, size = 3) +
  theme(axis.text.x = element_text(angle = 75, hjust=1,size = 9))
  
class_plt <- ggplot(live_vegetation, aes(fill=.pred_class, y = 1, x=area_class)) + 
   geom_bar(position="fill", stat="identity") +  
   theme(axis.text.x = element_text(angle = 75, hjust=1,size = 10)) +
  theme(legend.position = "bottom")

grid.arrange(sum_plt, class_plt, ncol=1,top = "Live Vegetation") 
```


### Sand
```{r fig.height = 8, fig.width = 9, fig.align = "center", dpi=350, echo=FALSE}
sand_data %>%
  mutate(area_class = cut(area, seq(0, max(area) + 100, 100), include.lowest = TRUE, right = TRUE ))  -> sand

sand %>%
  mutate(poly_id = paste(poly_num, quadrat, sep = "_")) %>%
  group_by(area_class, poly_id) %>%
  summarise(area = first(area), area_class = first(area_class)) %>%
  summarize(area = sum(area), poly_count = n() ) -> sand_summary 

sum_plt <-ggplot(sand_summary, aes(x = area_class, y = area, label = paste0(poly_count))) +
  geom_col()  +
  geom_text(nudge_y = 5000, size = 4) +
  theme(axis.text.x = element_text(angle = 75, hjust = 1, size = 9))
  
class_plt <- ggplot(sand, aes(fill=.pred_class, y = 1, x=area_class)) + 
   geom_bar(position="fill", stat="identity") +  
   theme(axis.text.x = element_text(angle = 75, hjust=1,size = 10)) +
  theme(legend.position = "bottom")

grid.arrange(sum_plt, class_plt, ncol=1,top = "sand") 
```

### Polygon Information
```{r}
val_pred %>%
  group_by(quadrat, poly_num, label) %>%
  summarise(
            count = n(),
            sand_count = length(.pred_class[ .pred_class == 'sand']),
            live_veg_count = length(.pred_class[ .pred_class == 'live vegetation']),
            dead_veg_count = length(.pred_class[ .pred_class == 'dead vegetation']),
            incorrect_count = length(.pred_class[ .pred_class != label ])
            ) %>%
  mutate(error = incorrect_count / count,
         live_veg_per = live_veg_count/count,
         dead_veg_per = dead_veg_count/count,
         sand_per = sand_count / count) -> counts 

counts %>% filter( poly_num == 13, quadrat == 39 )
counts
```

### Select Polygons where error rate is above threshold 
```{r}
counts %>%
  ungroup %>%
  filter(error > 0.50) %>%
  left_join(polys_sf %>%
               dplyr::select(-label), by = c('quadrat', 'poly_num')) %>%
  dplyr::select(-c(polys, imagePath, train_val, sand_count, incorrect_count, dead_veg_count, live_veg_count)) -> problem_polys

problem_polys %>%
  ungroup %>%
  pluck('area') %>%
  mean() -> mean

print('Average Area:')
mean* 0.0014

problem_polys %>% dplyr::select(label, quadrat, poly_num, area, error) %>% distinct() %>% arrange(desc(area))
```

### Find which class is the hardest
```{r}
counts %>%
  group_by(label) %>%
  summarise(mean = mean(error)) -> class_errors
class_errors
```

Dead vegetation was the hardest to classify with a 16 percent error rate

### Plot all the problem polygons
```{r}
problem_polys %>%
  dplyr::select(quadrat, poly_num)  %>%
  pmap(.f = plot_labeled) -> p
```

### Plot Labeled
```{r}
plot_labeled <- function(quadrat,
                         poly_num,
                         out_path = NULL,
                         maxpixels = Inf,
                         quadrats = sprintf("clean_data/quadrats/quadrat%02d/rgb.tif", seq(34,83)),
                         json_paths =sprintf("clean_data/quadrats/quadrat%02d/rgb.json", seq(34,83))){
  
  im_path <- quadrats[[quadrat]]
  json_path <- json_paths[[quadrat]]
  
  sf <- label_me_json_to_sf(json_path) %>%
    slice(poly_num) 
  
  buffer <- label_me_json_to_sf(json_path) %>% 
    slice(poly_num) %>% st_buffer(30)
  
  sf <- sf %>% rbind(buffer)
  
  label <- sf$label
  area <- sf$area
  
  ras <- brick(im_path)
  ras[is.na(ras)] <- 255
  crs(ras) <- NULL
  ras <- raster::crop(ras, extent(sf))
  
p <-ggplot() + 
    ggRGB(ras, r = 1, g = 2, b = 3, ggLayer = TRUE, maxpixels = maxpixels) +
        geom_sf(data = sf, fill = NA, expand = TRUE, size = 1.5) +
    theme_minimal() +
   ggtitle(paste('Quadrat:', quadrat, '\nPoly Number:', poly_num, '\nLabel:', label, '\nArea:', round(area) ))
  print(p)
  if(!is.null(out_path)){ggsave(out_path)}
}

plot_labeled(quadrat = 2, poly_num = 29, maxpixels = Inf)


```

Average area of polygons that misclassify pixels with an 

error rate > 50% is around 0.265 cm^2.

error rate > 25% is around 0.37 cm^2.

any error is around 2.956 

Of the polygons that had greater than 50% error the mean was .25 cm^2 is different than

Small polygons had the small error rate 

### Plotting
```{r}
quadrats <- sprintf("clean_data/quadrats/quadrat%02d", seq(34,83))
ras <- paste0(quadrats[[1]], "/rgb.tif") %>%
  brick()

p_orig <- ggRGB(ras, r = 1, g = 2, b = 3, maxpixels = 5e+05) + theme_minimal()
p_orig
```



### Polygon Area Distribution
```{r, echo = FALSE}
polys_sf %>%
  mutate(area_class = cut(area, seq(min(area), max(area), 10000))) %>%
  group_by(label) %>%
  summarise(total_area = sum(area))-> poly_area_summary

ggplot(poly_area_summary, mapping = aes(x = label, y = total_area, fill = label, label = paste( 'Area:', round(total_area)) )) +
  geom_text(nudge_y = 5e+4, size = 3.5) +
  geom_col() 
```

### Prediction Distribution
```{r, echo = FALSE}
val_pred %>%
  group_by(.pred_class) %>% 
  summarize(area = sum(area), n = n()) -> area_summary

ggplot(area_summary, mapping = aes(x = .pred_class, y = n, fill = .pred_class, label = paste("Count:",n))) +
  geom_col() +
  geom_text(nudge_y = 1e+4, size = 3.5) +
  scale_fill_manual(values=c("#b7a386", "#5a6233", "#5e4a31"))# +
  #scale_y_continuous(labels = scaleFUN)
```

### Compare Predictions to Actual

```{r fig.height = 5, fig.width = 8, fig.align = "center", dpi=350, echo=FALSE}
label_counts <- val_pred %>%
  group_by(label) %>%
  summarize(count = n()) %>%
  add_column( type = 'truth') %>%
  rename(class = label)

pred_counts <- val_pred %>%
  group_by(.pred_class) %>%
  summarise(count = n()) %>%
  add_column( type = 'prediction') %>%
  rename(class = .pred_class)

df <- label_counts %>%
  rbind(pred_counts) 

label_plt <-label_counts %>%
  ggplot(aes(x = class, y = count, label = count, fill = class)) +
  geom_col()  +
  geom_text(nudge_y = 7500, size = 5.5) +
  theme(axis.text.x = element_text(angle = 75, hjust=1,size = 12))+
  scale_fill_manual(values=c("#b7a386", "#5a6233", "#5e4a31")) +
  xlab("Label Counts")
  
class_plt <- pred_counts %>%
  ggplot(aes(x = class, y = count, label = count, fill = class)) +
  geom_col()  +
  geom_text(nudge_y = 7500, size = 5.5) +
  theme(axis.text.x = element_text(angle = 75, hjust=1,size = 12))+
  scale_fill_manual(values=c("#b7a386", "#5a6233", "#5e4a31")) +
  xlab("Prediction Counts")

grid.arrange(label_plt, class_plt, ncol=2,top = "Label Counts vs Prediction Counts") 

```
## Reading in data
<!-- ## Filter Live Veg -->
```{r, include = FALSE}
subset <- val_pred 
subset %>% filter(label == 'live vegetation') -> live_vegetation_data
subset %>% filter(label == 'dead vegetation') -> dead_vegetation_data
subset %>% filter(label == 'sand') -> sand_data
```

```{r, include = FALSE}
# subset_full <- val_pred_full 
# 
# subset_full %>% filter(label == 'live vegetation') -> live_vegetation_data_full
# subset_full %>% filter(label == 'dead vegetation') -> dead_vegetation_data_full
# subset_full %>% filter(label == 'sand') -> sand_data_full
```
<!-- ## Live Vegetation -->
<!-- ### Live Vegetation Exponential Cut -->
```{r fig.height = 10, fig.width = 9, fig.align = "center", dpi=350, echo=FALSE, include=FALSE ,eval = FALSE}
live_vegetation_data %>%
  mutate(area_class = cut(area, 10^seq(-2, 5, 0.25))) -> live_vegetation

#10^seq(-2, 5, 0.25))
live_vegetation %>% group_by(area_class) %>%
  summarize(area = sum(area), n = n()) %>%
  mutate(ave_area = area/n) -> live_vegetation_summary 

sum_plt <-  ggplot(live_vegetation_summary, aes(x = area_class, y = area, label = n)) +
  geom_col()  +
  geom_text(nudge_y = 1000000, size = 5) +
  theme(axis.text.x = element_text(angle = 75, hjust=1,size = 12))
  
class_plt <- ggplot(live_vegetation, aes(fill=.pred_class, y = 1, x=area_class)) + 
   geom_bar(position="fill", stat="identity") +  
   theme(axis.text.x = element_text(angle = 75, hjust=1,size = 12)) +
  theme(legend.position = "bottom")

grid.arrange(sum_plt, class_plt, ncol=1,top = "Live Vegetation Exponential") 
```

### Live Vegetation Regular Cut
```{r fig.height = 10, fig.width = 9, fig.align = "center", dpi=350, echo=FALSE}
live_vegetation_data %>%
  drop_na() %>%
  mutate(area_class = cut(area, seq(min(area),max(area),500))) -> live_vegetation

#10^seq(-2, 5, 0.25))

live_vegetation %>% group_by(area_class) %>%
  summarize(area = sum(area), n = n()) %>%
  mutate(ave_area = area/n) -> live_vegetation_summary 

sum_plt <-  ggplot(live_vegetation_summary, aes(x = area_class, y = area, label = area)) +
  geom_col()  +
  geom_text(nudge_y = 550000, size = 4) +
  theme(axis.text.x = element_text(angle = 75, hjust=1,size = 9))
  
class_plt <- ggplot(live_vegetation, aes(fill=.pred_class, y = 1, x=area_class)) + 
   geom_bar(position="fill", stat="identity") +  
   theme(axis.text.x = element_text(angle = 75, hjust=1,size = 10)) +
  theme(legend.position = "bottom")

grid.arrange(sum_plt, class_plt, ncol=1,top = "Live Vegetation Regular Cut") 
```

<!-- ## Sand -->

<!-- ### Sand Expotential Cut -->
```{r fig.height = 10, fig.width = 9, fig.align = "center", dpi=350, echo=FALSE, include=FALSE,eval = FALSE}
sand_data %>%
  mutate(area_class = cut(area, 10^seq(-2, 10, 0.25))) -> sand

sand %>% group_by(area_class) %>%
  summarize(area = sum(area), n = n()) %>%
  mutate(ave_area = area/n) -> sand_summary 

sand_sum_plt <-  ggplot(sand_summary, aes(x = area_class, y = area, label = n)) +
  geom_col()  +
  geom_text(nudge_y = 100000, size = 3) +
  theme(axis.text.x = element_text(angle = 75, hjust=1,size = 9))
  
sand_class_plt <- ggplot(sand, aes(fill=.pred_class, y = 1, x=area_class)) + 
   geom_bar(position="fill", stat="identity") +  
   theme(axis.text.x = element_text(angle = 75, hjust=1,size = 10)) +
  theme(legend.position = "bottom")

grid.arrange(sand_sum_plt, sand_class_plt, ncol=1,top = "Sand Exponential Cut") 
```


### Sand Regular Cut
```{r fig.height = 5, fig.width = 9, fig.align = "center", dpi=350, echo = FALSE}
sand_data %>%
  mutate(area_class = cut(area, seq( min(area), max(area), 25500), include.lowest = TRUE, right = TRUE)) -> sand

sand %>%
  mutate(poly_id = paste(poly_num, quadrat, sep = "_")) %>%
  group_by(area_class, poly_id) %>%
  summarise(area = first(area)) %>%
  summarize(n_pixels = n(), poly_count = n_distinct(poly_id)) -> sand_summary 

sand_sum_plt <-  ggplot(sand_summary, aes(x = area_class, y = n_pixels, label = poly_count)) +
  geom_col()  +
 # geom_text(nudge_y = 5e+8, size = 4) +
  theme(axis.text.x = element_text(angle = 75, hjust=1,size = 9))
  
sand_class_plt <- ggplot(sand, aes(fill=.pred_class, y = 1, x=area_class)) + 
   geom_bar(position="fill", stat="identity") +  
   theme(axis.text.x = element_text(angle = 75, hjust=1,size = 10)) +
  theme(legend.position = "bottom")

grid.arrange(sand_sum_plt, sand_class_plt, ncol=1,top = "Sand Regular Cut") 
```

<!-- ## Dead Vegetation -->

<!-- ### Dead Vegetation Expotential Cut -->
```{r fig.height = 10, fig.width = 9, fig.align = "center", dpi=350, echo=FALSE, include=FALSE,eval = FALSE}
dead_vegetation_data %>%
  mutate(area_class = cut(area, 10^seq(-2, 5, 0.25))) -> dead_vegetation

dead_vegetation %>% group_by(area_class) %>%
  summarize(area = sum(area), n = n()) %>%
  mutate(ave_area = area/n) -> dead_vegetation_summary 

sum_plt <-  ggplot(dead_vegetation_summary, aes(x = area_class, y = area, label = n)) +
  geom_col()  +
  geom_text(nudge_y = 500000, size = 4) +
  theme(axis.text.x = element_text(angle = 75, hjust=1,size = 9))
  
class_plt <- ggplot(dead_vegetation, aes(fill=.pred_class, y = 1, x=area_class)) + 
   geom_bar(position="fill", stat="identity") +  
   theme(axis.text.x = element_text(angle = 75, hjust=1,size = 10)) +
  theme(legend.position = "bottom")

grid.arrange(sum_plt, class_plt, ncol=1,top = "Dead Vegetation Expotential") 
```

### Dead Vegetation Regular Cut

Probably a pattern like we expected. 


